---
---

## Categorical data

```{r}
default.stringsAsFactors()
```

😲 this setting is new in R version 4 !!

We used read_csv so it was that way anyway, but there are situations where you may want to use factors, especially for MODELING and PLOTS. essentially when you want character vectors to not be alphabetical. The `forcats` package is intended to make this easier.

===

The column with the species common names is a character vector

```{r}
str(pg_df$common)
```

Plots will be sorted alphabetically, like the default ordering for factor levels in base R.

```{r}
as.factor(pg_df$common) %>% levels()
```

```{r, fig.height = 3, dpi=100}
pg_df %>% 
  ggplot(aes(x = common)) +
  geom_bar()
```

===

Forcats `as_factor` uses the order in which the levels appear in your data

```{r}
as_factor(pg_df$common) %>% levels()
```

Use this to make common into a factor. Notice the ordering of the bars now.

```{r, fig.height = 3, dpi=100}
pg_df <- pg_df %>% 
  mutate(common = as_factor(common))

pg_df %>% 
  ggplot(aes(x = common)) +
  geom_bar()
```


===

Use `fct_infreq` to instead order the levels by how frequently they appear in the dataset.

```{r, fig.height = 3, dpi=100}
pg_df %>% 
  mutate(common = fct_infreq(common)) %>%
  ggplot(aes(x = common)) +
  geom_bar()
```

===

Or more generally, `fct_reorder` to use a function(`.fun`) of a different variable (`.x`). In this case let's order **common** by each level's median `flipper_length_mm`. Additional arguments to the specified function can also be included. 

```{r, warning=FALSE, fig.height = 3, dpi=100}
pg_df %>% 
  mutate(common = fct_reorder(common, 
    .x = flipper_length_mm, .fun = median, na.rm = TRUE)) %>%
  ggplot(aes(x = common, y = flipper_length_mm)) +
  geom_boxplot()
```

===

## All combinations of variables

To create a table with combinations of variables, use `expand` from the tidyr package. For example:

```{r}
pg_df %>% expand(island, sex, common)
```

To only include the combinations of variables that do appear in your data, put the column names inside of the `nesting` helper function.

```{r, handout = 0}
pars <- pg_df %>% expand(nesting(island, sex, common))
```
```{r}
str(pars)
```




