---
---

## Categorical data

R has a special data type for handling categorial data called factors. Because they had a reputation for causing trouble... 

```{r}
default.stringsAsFactors()
```

😲 this setting is new in R version 4 !!

Although `read_csv` had always defaulted strings to character data, there are situations where you may want to use factors, especially for modeling and plots -- essentially whenever you want character vectors to not be alphabetical. [forcats](){:.rlib} is intended to make this easier.

===

Species common names are in a character vector. 

```{r}
class(pg_df$common)
```

Controlling the order in which the 3 species appear in a plot is done with factors. If the data are kept as characters, plots will be sorted alphabetically (i.e. the default ordering for factor levels in base R).

```{r}
as.factor(pg_df$common) %>% levels()
```

```{r, fig.height = 3, dpi=100, handout = 0}
pg_df %>% 
  ggplot(aes(x = common)) +
  geom_bar()
```

===

`forcats::as_factor` uses instead the order in which the levels appear in your data:

```{r}
as_factor(pg_df$common) %>% levels()
```

Convert `common` into a factor and remake the same bar plot. Notice the ordering of the bars compared to the previous graph. 

```{r, fig.height = 3, dpi=100, handout = 0}
pg_df <- pg_df %>% 
  mutate(common = as_factor(common))

pg_df %>% 
  ggplot(aes(x = common)) +
  geom_bar()
```


===

Use `fct_infreq` to order the levels by how frequently they appear in the dataset.

```{r, fig.height = 3, dpi=100, handout = 0}
pg_df %>% 
  mutate(common = fct_infreq(common)) %>%
  ggplot(aes(x = common)) +
  geom_bar()
```

===

Or more generally, `fct_reorder` to use a function(`.fun`) of a different variable (`.x`). Order **common** by each level's median `flipper_length_mm`:

```{r, warning=FALSE, fig.height = 3, dpi=100, handout=0}
pg_df %>% 
  mutate(common = fct_reorder(common, 
    .x = flipper_length_mm, .fun = median, na.rm = TRUE)) %>%
  ggplot(aes(x = common, y = flipper_length_mm)) +
  geom_boxplot()
```

===

## All combinations of variables

To create a table with combinations of variables, use `expand` from the tidyr package. For example:

```{r}
pg_df %>% expand(island, sex, common)
```

To only include the combinations of variables that do appear in your data, put the column names inside of the `nesting` helper function.

```{r, handout = 0}
pars <- pg_df %>% expand(nesting(island, sex, common))
```
```{r}
str(pars)
```




