---
editor_options: 
  chunk_output_type: console
---

## Tidy evaluation

Another core tenet of the tidyverse is *tidy evaluation* via data masking, which facilitates:

- using variable/column names without having to refer to data frame name
- identifying variables based on their position, name, or type

In base R there is similar functionality using `attach(pg_df)` or `data = pg_df` arguments in functions like `lm`. Tidyverse functions take those ideas and make them consistent and operational in a reliable manner beyond usage in interactive programming.
{:.notes}

===

We will rename the columns of our data frame to remove spaces and punctuation.

```{r}
names(pg_df)
```


===

We could rename each individually using `dplyr`'s `rename`, but `rename_with` is a handy shortcut if you want to do the same thing to every column name:

```
rename_with(.data = pg_df, .fn = ...)
```

We will use functions in the core tidyverse `stringr` package to manipulate the column name strings. 

===

## Pattern replacement 

Replaces spaces with underscore characters ("_") either with exact pattern or [regular expressions](https://stringr.tidyverse.org/articles/regular-expressions.html). Test out function on one column name:

replace vs replace all

```{r, eval=FALSE}
str_replace_all("Body Mass (g)", pattern = " ", replacement = "_")
str_replace_all("Body Mass (g)", pattern = "[:space:]", replacement = "_")
```

===

Similarly, can remove all punctuation using the regex pattern `[:punct:]`

```{r, eval=FALSE}
str_replace_all("Delta 15 N (o/oo)", pattern = "[:punct:]", replacement = "")
```

Combine both transformations and add a new one (convert to lower case) in one fell swoop:

point out function syntax again

```{r, handout = 0}
pg_df <- pg_df %>%
  rename_with(~str_replace_all(.x, "[:punct:]", "")) %>%
  rename_with(~str_replace_all(.x, "[:space:]", "_")) %>%
  rename_with(~str_to_lower(.x))
```

Check out the new column names:

```{r}
names(pg_df)
```

Much improved!!

===

## Format dates

Now let's work on putting the date information from the filenames into a properly formatted column of dates. Look at a sample of the data to evaluate: Is there a systematic way to extract the necessary information?

```{r}
sample(pg_df$filename, 5)
```

===

Every file name is the same besides the date. We can remove the patterns that match "data/penguins/penguins_nesting-" or (`|`) ".csv" using `str_remove_all`. Save a new vector so we can focus on the date conversion. 

```{r, handout = 0}
egg_dates <- str_remove_all(string = pg_df$filename, 
    pattern =  "(data/penguins/penguins_nesting-)|(.csv)")
```

===

`lubridate` is a tidyverse package to facilitate working with date formats. Can `as_date` automatically interpret "December-1-2009" as a date?

```{r, eval=FALSE, results='hide', handout = 0}
library(lubridate)
egg_dates[1] %>% as_date()
```


===

As with the base `as.Date` function, unless dates are already specified using the [ISO 8601](https://en.wikipedia.org/wiki/ISO_8601) YYYY-MM-DD representation, date conversion with `lubridate::as_date` requires a character string defining the date using codes from the POSIX standard:

| code    | meaning          |
|---------+------------------|
| `%b`    |  Abbreviated month name in current locale   |
| `%B`    |  Full month name in current locale          |
| `%d`    |  Day of month as decimal number (01-31)     |
| `%j`    |  Day of year as decimal number (001-366)    |
| `%y`    |  Year without century (00-99)               |
| `%Y`    |  Year with century                          |

Note that interpretation of many POSIX codes depend on settings of your operating system, such as month or weekday names in different languages, whether weeks start on Sunday or Monday, or the default century for `%y`. Read more about format specification in the **Details** section of the base `strptime` function. 

```{r, eval = FALSE}
?strptime
```

===

Supply the pattern of the date format as a character string using the appropriate codes:

```{r, handout = 0}
egg_dates[1] %>% as_date(format = "%B-%d-%Y")
```

As long as the components of a date are in a consistent order, the `guess_formats` function can help determine the appropriate pattern, by supplying the order of **m**onth, **d**ay, and **y**ear. 

```{r, results='hide', handout = 0}
egg_dates[1] %>% guess_formats(orders = "mdy")
```

===

Just by knowing the order that months, days, and years appear, we can actually circumvent all of those previous steps to find and convert the date information right from the full filename using the function `mdy` 

```{r}
pg_df$filename[1] %>% mdy()
```

Use this function to create a new column with sampling dates.

```{r, handout = 0}
pg_df <- pg_df %>% mutate(date = mdy(filename))
```



