---
---

## Tidy evaluation

Another core tenet of the tidyverse is *tidy evaluation* via data masking, which facilitates:

- using variable/column names without having to refer to data frame name
- identifying variables based on their position, name, or type

In base R there is similar functionality using `attach(pg_df)` or `data = pg_df` arguments in functions like `lm`. Tidyverse functions take those ideas and make them consistent and operational in a reliable manner beyond usage in interactive programming.
{:.notes}

===

We will rename the columns of our data frame to remove spaces and punctuation.

```{r}
names(pg_df)
```


===

We could rename each individually using `dplyr`'s `rename`, but `rename_with` is a handy shortcut if you want to do the same thing to every column name:

```
rename_with(.data = pg_df, .fn = ...)
```

We will use functions in the core tidyverse `stringr` package to manipulate the column name strings. 

===

## Pattern replacement 

Replaces spaces with underscore characters ("_") either with exact pattern or [regular expressions](https://stringr.tidyverse.org/articles/regular-expressions.html). Test out function on one column name:

```{r, eval=FALSE}
str_replace_all("Body Mass (g)", pattern = " ", replacement = "_")
str_replace_all("Body Mass (g)", pattern = "[:space:]", replacement = "_")
```

===

Similarly, can remove all punctuation using the regex pattern `[:punct:]`

```{r, eval=FALSE}
str_replace_all("Delta 15 N (o/oo)", pattern = "[:punct:]", replacement = "")
```

Combine both transformations and add a new one (convert to lower case) in one fell swoop:

```{r, handout = 0}
pg_df <- pg_df %>%
  rename_with(~str_replace_all(.x, "[:punct:]", "")) %>%
  rename_with(~str_replace_all(.x, "[:space:]", "_")) %>%
  rename_with(~str_to_lower(.x))
```

Check out the new column names:

```{r}
names(pg_df)
```

Much improved!!

===

## Format dates

Now let's work on putting the date information from the filenames into a properly formatted column of dates. Look at a sample of the data to evaluate: Is there a systematic way to extract the necessary information?

```{r}
sample(pg_df$filename, 5)
```

===

Every file name is the same besides the date, so we can remove those exact characters. Save to a vector. 

```{r, handout = 0}
egg_dates <- str_remove_all(string = pg_df$filename, 
    pattern =  "(data/penguins_nesting-)|(.csv)")
```

===

Can these be automatically converted to date format?

```{r, eval=FALSE}
as.Date(egg_dates)
```


===

Lubridate can make some smart guesses about date formats with a little bit of help: the order **m**onth, **d**ay, **y*ear. 

```{r, handout = 0}
library(lubridate)
mdy(egg_dates)
```

===

The POSIX date standard specifies codes for various components of dates and times, which can be used for the `format` argument in `as_date`

```{r, results='hide'}
guess_formats("December-1-2009", orders = "mdy")
as_date(egg_dates, format = "%B-%d-%Y")
```

===

Combine the string manipulation and date conversion: 

```{r, handout = 0}
pg_df <- pg_df %>%
  mutate(egg_date = str_remove_all(filename,
          "(data/penguins_nesting-)|(.csv)")) %>%
  mutate(egg_date = mdy(egg_date))
```



